--- 
title: "splithalf_documentation"
author: "Sam Parsons"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: sdparsons/splithalf_documentation
description: "This is the documentation for the splithalf package."
---

# Start-up notes

## Instalation 

The **splithalf** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("splithalf")
# or the development version
# devtools::install_github("sdparsons/splithalf")
```

also note that the package requires the following; _tidyr_, _dplyr_, _Rcpp_, _robustbase_. You can load them with the following

```{r eval = FALSE}
library("tidyr")
library("dplyr")
library("Rcpp")
library("robustbase")
```


## Version note

The current version of **splithalf** is 0.5.1 [unofficial version name: Fight Milk]

A huge part of developing this latest version has been condensing the many splithalf functions into a single one. In addition to this, I have adapted the underlying code using the amazing Rcpp package [@R-Rcpp], enabling me to input some C++ code. While my C++ code is rudimentary, the _splithalf_ function now runs around 20x faster than before. 


## Citing the package

Citing packages is one way for developers to gain some recognition for the time spent maintaining the work. I would like to keep track of how the package is used so that I can solicit feedback and improve the package more generally.  This would also help me track the uptake of reporting measurement reliability over time. 
Please use the following reference: Parsons, S. (2017). splithalf: robust estimates of split half reliability
  (Version 2). figshare. https://doi.org/10.6084/m9.figshare.5559175.v2


## A note on terminology used in this document

It is important that we have a similar understanding of the terminology I use in the package and documentation. Each is also discussed in reference to the functions later in the documentation. 

* Trial – whatever happens in this task, e.g. a stimuli is presented. Importantly, participants give one response per trial
* Trial type – often trials can be split into different trial types (e.g. to compare congruent and incongruent trials)
* Condition - this might be different blocks of trials, or something to be assessed separately within the functions. e.g. a task might have a block of 'positive' trials and a block of 'negative' trials.
* Datatype - I use this to refer to the outcome of interest. specifically whether one is interested in average response times or accuracy rates
* Score - I use score to indicate how the final outcome is measured; e.g. the average RT, or the difference between two average RTs, or even the difference between two differences between two RTs (yes, the final one is confusing)
